Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-04-07 Ben_Manuscripts/hdphmm/Outline.tex
Reworking outline with radial clusters

  Solutes in this system show a wide range of behavior influenced by the 
  heterogeneity of the membrane's nanostructure as well as the interactions 
  between monomer and solute chemical functionality. 
  \begin{itemize}  
  	\item Our application of the IHMM does a good job of identifying and
  	distinguishing these behaviors.   	
  	\item Clustering the behaviors helps us to tell a coherent story which
  	emphasizes those interactions which most heavily influence solute transport.
  	\item This analysis approach helps us to break down the very complex behavior
  	exhibited by solutes over long MD trajectories into discrete categories.
  \end{itemize}

@@

  %BJC3: obviously can use a better thesis.
  We showcase this modeling approach by example. 
  \begin{itemize}
    \item In Figure~\ref{fig:mechanism_map}, we plot a representative 2D methanol 
    trajectory in cylindrical coordinates.
    \item The coordinates are colored according to the cluster with which methanol's
    dynamics are most consistent, which implies that this particular methanol 
    trajctory exhibits 6 distinct dynamical behaviors. 
    \item The solute timeseries are coupled with color coded bars which describe the
    solute's local environment and physical interactions that it undergoes at that
    time frame.
    %MRS: maybe number (1-4) or label (a-d) the rows as well. 
    %BJC4: yeah, meant to do that. As it stands, I'm referring to 1 through 4 as bottom through top.
    \item From the top, they describe whether the solute is associated with sodium,
    whether the solute is hydrogen bonding with the monomer as well as what part of
    the monomer it is hydrogen bonding to, the local number density surrounding the
    solute and the radial location of the solute with respect to the nearest pore
    center (this information is redundant but can be a useful visual aid). %BJC3: still debating removing that bar
  \end{itemize}

  %BJC3: I still need labels for this plot
  %BJC4: Also, for some reason, when saving to a PDF, the colorbars are faded. I'll fix this eventually.
  %MRS5: explicitly state how many states there are, and their colors.
  \begin{figure}
  \includegraphics[width=\textwidth]{mechanism_map.pdf}
  \caption{We can learn a significant amount of detail about solute motion by viewing 
  solute trajectories color-coded according to their clustered dynamical behavior alongside
  plots of the dominant molecular interactions and descriptions of the membrane's local 
  environment. In the plot above, we show the radial and axial coordinates of a 
  %MRS5:
  %chosen
  single
  methanol molecule over the course of the equilibrated portion of its MD simulation.
  Disjoint segments of the same color imply the solute behaves similarly at different 
  points of the trajectory because their VAR parameters %
  %MRS5:
  %are represented by 
  place them in
  the same 
  %BJC4: eventually I'll label the bars and probably put a picture of the monomer with different oxygens labeled
  cluster. 
  %MRS5:
  %Above 
  On the top of
the plot are bars which represent different physical interactions and
  the membrane's local environment as a function of time. The top bar is colored green 
  at time points where solutes are associated with sodium. Methanol associates
  with sodium ions relatively rarely. The next bar down is colored when hydrogen bonds exist. Blue slivers
  correspond to hydrogen bonds donated to the monomer's carboxylate head group. Green slivers
  corrrespond to  hydrogen bonds donated to the ether linkages between the head groups
  and tails. Red slivers correspond to hydrogen bonds donated to the oxygens on the ends
  of the monomer tails. Methanol exhibits significant hydrogen bonding with all parts of
  the monomer. Which part of the monomer methanol is hydrogen bonds to is heavily correlated
  to the solute's $r$ coordinate. The bar with the blue gradient represents the local 
  membrane density based on the solute's ($r$, $z$) coordinate. Darker shades of blue
  %BJC4: eventually I want some kind of color bar legend
  correspond to higher membrane densities. The density appears to be higher when methanol
  motion appears restricted (e.g. from 75 to 550 ns). It is low in areas where methanol's
  position fluctuates significantly (e.g. from 550 to 700 ns). Finally, the bottom bar
  simply color-codes the solute's radial distance from the pore center. This information
  is the same as shown by the radial coordinate but can be a helpful visual aid when
  studying the other bars.}\label{fig:mechanism_map}
  \end{figure}
  %MRS5: hmm, one sign that someting might be nonideal is the clustering in the light blue state (assuming 1 light blue state), looking at the 75 nm to 550 nm segment and the 900 - 1050 nm segment, the r is quite different (1 vs 3), and the h-bonding patterns are very different (Green vs. none). So it's not entirely clear these should necessarily be the same state. Thoughts?
  One can gain significant chemical intuition just by studying plots like
  Figure~\ref{fig:mechanism_map}.
  \begin{itemize}
    \item It is clear that, for methanol, association with sodium is a much rarer
    interaction than hydrogen bonding with the LLC monomers.
    \item Methanol hydrogen bonds to all regions of the monomers dependent on
    its radial position.
    \item Its fluctuations tend to be smallest when hydrogen bonded or in areas
    of high local number density. 
    \item For example, from approximately 75 to 525 ns, methanol appears to be trapped
    in a high density region of the tails.
    \item In the time that follows, it enters a region of relatively low density
    where fluctuations are quite large. 
    \item During this time period, there is intermittent, short-lived hydrogen 
    bonding with the oxygen atoms at the tail ends, as suggested by the frequent 
    state changes during that time period and red slivers in the bar above.
  \end{itemize}

@@

  We may learn the most by studying the dynamical modes common to the majority of 
  the solute trajectories.
  \begin{itemize}
    \item Using the IHMM, we identified 13 total dynamical modes exhibited by 
    methanol. 
    \item Of those, 4 appear in at least 50 \% of the methanol trajectories, and
    2 of them appear in over 90 \% of the methanol trajectories.
    \item In Figure~\ref{fig:common_states_MET_lines}, we plot representative dynamics
    of each of the 4 prevelant states.
    \item Based on their self-transition probabilities, we estimated the average 
    time spent by each solute in each state (Figure~\ref{fig:dwell_times_MET}).
    \item States 2 and 4 have long dwell times relative to states 1 and 3 because
    they have high self-transition probabilities.
    \item States 1 and 3 show large fluctuations relative to the other states owing
    to larger covariances in both the axial and radial dimensions. (Figure~\ref{fig:A_sigma_scatter_MET})
    \item States 1 and 2 have lower autoregressive coefficients relative to states
    3 and 4.
    \item All of these states are present in Figure~\ref{fig:mechanism_map} and colored
    accordingly.
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  % BJC: I'll bootstrap some errorbars on this by drawing the transition probabilities 
  % from a dirichlet distribution.
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}
  \caption{}\label{fig:dwell_times_MET}
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \end{subfigure}
  \caption{We can learn the most about methanol's behavior by studying the dynamics
  of the states common to the majority of methanol trajectories. In (a), we show representative
  dynamics of four states that are common to more than half of the solute trajectories. The
  percentage of trajectories in which they appear are listed between the $r$ and $z$ plots. 
  All four of these states are exhibited in the trajectory shown in Figure~\ref{fig:mechanism_map}
  and are colored to match for easy visual comparison. All time series have a mean of zero and are
  %MRS5: state how much shifted - or maybe plot a dotted line at the zero of each plot?
  shifted for the purpose of visualizing them. (b) Using the transition probabilities (p), 
  we estimated the average time spent within each of the states using 1000 simulations of a Bernoulli 
  process where we continued to draw until transitioning out of the state with probability 1 - p.
  On average, methanol stay in states 2 and 4 for much longer periods of time than states 1 and 3.
  This is qualitatively evident in Figure~\ref{fig:mechanism_map}. (c) We can use the understanding
  from (a) and (b) in order to relate the state parameters to solute motion. States 1 and 3 have
  large covariances relative to states 2 and 4. Paired with their short dwell times, it is reasonable
  to hypothesize that states 1 and 3 contribute to hopping behavior while states 2 and 4 
  correspond to trapped solute behavior. All states have positive autoregressive coefficients (A) in 
  each dimension which implies motion that is positively correlated to its previous step. States 3 and
  4 have higher autoregressive coefficients than 1 and 2. This implies that solutes in states 3 and 
  4 tend to wander more from their mean position in each dimension. This appears to be the case when
  comparing the behavior of states 2 and 4 in Figure~\ref{fig:mechanism_map}.}\label{fig:common_states_MET}
  \end{figure}

@@

  %BJC3: this figure is not even close to done
  %BJC3: I've tried a couple different things to look at the hbonds. The other was 
  % how often they are hydrogen bonded in general versus not hydrogen bond. They all
  % hydrogen bond about the same amount. So I think hbond lifetimes may be the key. 
  % I need to implement this first to test that hypothesis.
  %BJC4: I didn't put sodium ion assocaition in this because it's infrequent. 
  %MRS5: is it correlated with specific states?
  %BJC4: If I compare to acetic acid or urea, I could add it back. I was thinking of splitting
  % the pi chart in half with slices corresponding to hbonds on one side and association on the other
  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{hbond_pi_charts.pdf}
  \caption{With our clustering approach, methanol exhibits similar state behaviors
  %MRS5: So this is a little counterintuitive, in that the separate states have similar behavior.
  %MRS5: Are there charts that show what is _different_ about the states as well?  The previous figure
  %      Shows that they are dynamically different, but this structural pie charge says the existance
  %      of H-bonding is not the reason . . . 
  in all regions of the membrane. In all of the common states described in 
  Figure~\ref{fig:common_states_MET}, methanol, at some point, hydrogen bonds with 
  each region of the LLC monomer to an appreciable degree. However, methanol frequently
  is not hydrogen bonded at all.
  %BJC4: hypothesis yet to be tested. Somehow, I'll incorporate lifetimes into this plot
  %MRS4: lifetimes would be good!
  We can learn more by looking at the hydrogen bond lifetime. In states 2 and 4, solutes
  stay hydrogen bonded for longer periods of time.
  %MRS5: OK, here is some physical insight - local density is what determines trapping. Emphasize the physical mechanisms (wat does and doesn't come out physically because of these analyses) like this more explicitly.
  %MRS5: can you make other graphs like the H-bond graph? like histogram/pie chart of local density
  %MRS5: as a function of state?  What about r as a function of state?  what other ways are there to show
  %MRS5: the characteristics of the states?
  These longer lifetimes exist because, when hydrogen bonded, the local density is higher, preventing the 
  bond from breaking so easily. In states 1 and 3, hbonds are fleeting interactions because
  the surrounding density is low, allowing solutes to move more freely.
  }\label{fig:hbond_pichart}
  \end{figure}

>>>>

  Solutes in this system show a wide range of behavior influenced by the 
  heterogeneity of the membrane's nanostructure as well as the interactions 
  between monomer and solute chemical functionality. 
  \begin{itemize}  
  	\item Our application of the IHMM does a good job of identifying and
  	distinguishing these behaviors.   	
  	\item The following approach to analysis demonstrates how one can break 
  	down the very complex behavior exhibited by solutes over long MD trajectories 
  	into discrete categories.
%  	\item Clustering the behaviors helps us to tell a coherent story which
%  	emphasizes those interactions which most heavily influence solute transport.
	\item Further, we can use the IHMM in order to generate stochastic trajectories,
	with dynamics similar to MD, so that we can estimate macroscopic transport
	properties.
  \end{itemize}

@@

  We applied the IHMM to all trajectories of each of the four solutes studied.
  \begin{itemize}
    \item Based on our choice of clustering, we identified over 30 distinct dynamical
    modes for each solute.
    \item Many of these modes are sampled very infrequently and thus do not merit significant study.
    \item In this section, we explore a subset of this data, focusing on the states
    and dominant interactions most common to solute trajectories. 
  \end{itemize}

  \noindent Methanol exhbits a range of dynamic behavior.
  \begin{itemize}
    \item In Figure~\ref{fig:mechanism_map}, we plot a representative 2D methanol 
    trajectory in cylindrical coordinates.
    \item The coordinates are colored according to the cluster with which methanol's
    dynamics are most consistent, which implies that this particular methanol 
    trajctory exhibits 8 distinct dynamical behaviors (in order of their first appearance,
    they are colored dark yellow, red, light blue, light green, orange, blue, teal and lime green). 
    \item The solute timeseries are coupled with color coded bars which describe the
    solute's local environment and physical interactions that it undergoes at that
    trajectory frame.
    \item From the top, they describe whether the solute is associated with sodium,
    whether the solute is hydrogen bonding with the monomer as well as to what part of
    the monomer it is hydrogen bonding, and finally, the local number density 
    surrounding the solute. 
    %and the radial location of the solute with respect to the nearest pore
    %center (this information is redundant but can be a useful visual aid).
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.75\textwidth}
  \includegraphics[width=\textwidth]{mechanism_map.png}
  \caption{}
  \end{subfigure}
  \begin{subfigure}{0.24\textwidth}
  \includegraphics[width=\textwidth]{monomer_oxygens.png}
  \caption{}
  \end{subfigure}
  \caption{(a) We can learn a significant amount of detail about solute motion by viewing 
  solute trajectories color-coded according to their clustered dynamical behavior alongside
  plots of the dominant molecular interactions and descriptions of the membrane's local 
  environment. In the plot above, we show the radial and axial coordinates of a single
  methanol molecule over the course of the equilibrated portion of its MD simulation.
  Disjoint segments of the same color imply the solute behaves similarly at different 
  points of the trajectory because their VAR parameters place them in the same cluster. 
  There are eight distinct dynamical modes exhibited by this trajectory. In order
  of their first appearance, they are colored dark yellow, red, light blue, light green,
  orange, blue, teal and lime green. On the top of the plot are bars which represent different
  physical interactions and the membrane's local environment as a function of time. The top 
  bar is colored green at time points where solutes are associated with sodium. Methanol 
  associates with sodium ions relatively infrequently. The next bar down is colored when hydrogen
  bonds exist. The colors match the enlarged oxygen atoms highlighted in (b). Blue slivers
  correspond to hydrogen bonds donated to the monomer's carboxylate head group. Green slivers
  corrrespond to  hydrogen bonds donated to the ether linkages between the head groups
  and tails. Red slivers correspond to hydrogen bonds donated to the oxygens on the ends
  of the monomer tails. Methanol exhibits significant hydrogen bonding with all parts of
  the monomer. Which part of the monomer methanol hydrogen bonds to is heavily correlated
  to the solute's $r$ coordinate. Finally, the bar with the blue gradient represents the local 
  membrane density based on the solute's ($r$, $z$) coordinate. Darker shades of blue
  correspond to higher membrane densities. The density appears to be higher when methanol
  motion is restricted (e.g. from 75 to 550 ns). It is relatively low in areas where methanol's
  position fluctuates significantly (e.g. from 550 to 700 ns). 
  %Finally, the bottom bar
  %simply color-codes the solute's radial distance from the pore center. This information
  %is the same as shown by the radial coordinate but can be a helpful visual aid when
  %studying the other bars.
  }\label{fig:mechanism_map}
  \end{figure}

  \noindent One can gain useful chemical intuition just by studying plots like
  Figure~\ref{fig:mechanism_map}.
  \begin{itemize}
    \item It is clear that, for methanol, association with sodium is a much rarer
    interaction than hydrogen bonding with the LLC monomers.
    \item Methanol hydrogen bonds to all regions of the monomers dependent on
    its radial position.
    \item Its fluctuations tend to be smallest when hydrogen bonded or in areas
    of high local number density. 
    \item For example, from approximately 75 to 525 ns, methanol appears to be trapped
    in a high density region of the tails.
    \item In the time that follows, it enters a region of relatively low density
    where fluctuations are quite large. 
    \item During this time period, there is intermittent, short-lived hydrogen 
    bonding with the oxygen atoms at the tail ends, as suggested by the frequent 
    state changes during that time period and red slivers in the bar above.
    \item Signficant hops in the $z$ direction appear to be described by the blue
    state.
    \item After entering the blue state ca. 850 ns, methanol appears to jump
    signficantly in the $z$ direction from a hydrogen bonded state with the ether 
    linkages to a hydrogen bonded state with a head group in a monomer below it.
  \end{itemize}

@@

  \noindent It is clear that there is a wealth of information in these plots. 
  \begin{itemize}
    \item One can analyze each segment sequentially and form a clear picture
    of this instance of methanol's behavior without the need to visulize
    the traejectory with software like VMD.
    \item However, a more general understanding of methanol's average behavior
    is necessary in order to formulate LLC membrane design principles.
  \end{itemize}

  %BJC: I'm not sure whether to spell out the numbers or not. 'six state' versus '6 states'; 'eight out of 24' or '8 out of 24'
  % 'State one or State 1' 
  \noindent We may learn the most by studying the dynamical modes common to the majority of 
  the solute trajectories.
  \begin{itemize}
    \item Using the IHMM, we identified 33 total dynamical modes exhibited by 
    methanol. 
    \item Of those, only six states appear in a third (8 out of 24) or more methanol 
    trajectories, and 3 of them appear in over 50 \% of the methanol trajectories.
    \item 12 of the detected states are shared by two or less methanol trajectories.  % 23 in 3 or less -- not sure which is better to say
    \item In Figure~\ref{fig:common_states_MET_lines}, we plot representative dynamics
    of each of the six states which we determined to be prevelant.
    \item We also chose to study a seventh state, which only appears in three out of
    24 trajectories, because methanol experiences very long periods of entrapment
    while in this state. Long trapping events, even if they are relatively rare, will
    have a significant influence on solute flux.
    \item No trajectory contains all seven states. However, the trajectory in 
    Figure~\ref{fig:mechanism_map} exhibits all modes except states 6 and 7.
    \item The states in Figure~\ref{fig:common_states_MET} are colored to match
    states that also appear in Figure~\ref{fig:mechanism_map}.
  \end{itemize}

  \noindent We can begin to dissect the dynamics of each state by studying their parameters.
  \begin{itemize}
    \item States 1, 3 and 5 have radial means near 1 nm or less (see y axis of $r$ direction plot in
    Figure~\ref{fig:common_states_MET_lines}), which suggest that they occur
    close to the pore centers. 
    \item States 2, 4, 6 and 7 occur well within the tail region of the membrane.
    \item States 1 and 2 have higher covariances in both the axial and radial
    directions compared to all other states.
    \item Covariance in each dimension appears to have a strong correlation.
    \item States 3 and 5 as well as states 4, 6 and 7 are primarily distinguished based
    on their autoregressive coefficients, A. 
    \item Higher values of A generally lead to more wandering about the state mean.
    \item One could argue that the parameters of states 3--7 are actually very close which
    suggests that solutes might behave alike in topologically distinct regions of the 
    membrane.
  \end{itemize}

  Despite some similarity in their VAR parameters, the length of time spent in states 3--7 
  appears to be dependent on the solute's radial location.
  \begin{itemize}
    \item Based on the self-transition probabilities, we estimated the average time
    spent by each solute in each state (Figure~\ref{fig:dwell_times_MET}).
    \item States 3, 5 and 7 have long dwell times relative to the rest because
    they have high self-transition probabilities.
    \item Somewhat surprisingly, both states 3 and 5 correspond to solutes located close to
    a pore center, where one might expect solute motion to be less restricted.
    \item Although solutes in states 4 and 6 are in the viscous tail region, their
    dwell times are only intermediate. This could be due to imhomogeneity in the
    tail region which causes transitions between many different trapped states but
    with litle net effect on solute MSD. 
    \item However, there are instances of states within the tail region with dwell 
    times that exceed 100--200 ns, as shown by our inclusion of state 7.
%    \item These long dwell times have a significant effect on dynamics and are
%    worth studying.
    \item There are no other states with dwell times longer than state 5 in the pore 
    region.
  \end{itemize}

  \noindent States 1 and 2 exhibit the shortest dwell times of the states studied.
  \begin{itemize}
      \item This is consistent with the trajectory in Figure~\ref{fig:mechanism_map}.
      \item State 2 is the shortest lived state, appearing to bridge other tail region
      states. 
      \item State 1 appears to represent a significant hopping state. 
      \item It has a high covariance and occurs near the pore center, where there is
      less obstruction to solute motion.
      \item It's dwell time of ca. 8 ns, suggests that solutes stay in this highly
      mobile state for long enough to displace signficantly.
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.6\textwidth}
  \includegraphics[width=\textwidth]{common_states_MET.pdf}
  \caption{}\label{fig:common_states_MET_lines}
  \end{subfigure}
  \begin{subfigure}{0.35\textwidth}
  % BJC: I'll bootstrap some errorbars on this by drawing the transition probabilities 
  % from a dirichlet distribution.
  \includegraphics[width=\textwidth]{A_sigma_scatter_MET.pdf}
  \caption{}\label{fig:A_sigma_scatter_MET}
  \includegraphics[width=\textwidth]{dwell_times_MET.pdf}  %BJC5: maybe better as 95 % confidence rather than mean
  \caption{}\label{fig:dwell_times_MET}
  \end{subfigure}
  \caption{We can learn the most about methanol's behavior by studying the dynamics
  of the states common to many of the methanol trajectories. In (a), we show representative
  dynamics of six states that are common to more than one third of the solute trajectories. 
  The $y$-axis labels of the $r$ coordinate plot specify the radial mean of each clustered 
  state. The percentage of trajectories in which the states appear are labeled to the right of
  the $r$ plots. States 1--5 appear in the trajectory shown in Figure~\ref{fig:mechanism_map}
  and are colored to match for easy visual comparison. All time series have a mean of zero and are
  shifted to put the mean in line with the $y$-axis labels for the purpose of visualizing them. 
  (b) We can begin to understand the solute behaviors demonstrated in (a) by understanding their VAR 
  parameters. States 1 and 2 have higher covariances in both the axial and radial directions compared
  to all other states. This should lead to signficant solute displacement. Covariance in each dimension
  appears to have a strong correlation. States 3--6 all have relatively low covariances, which may
  be characteristic of trapped states. They are primarily distinguished based on the magnitude of their 
  autoregressive coefficients, A, and radial means. Higher values of A imply motion which wanders 
  more about the mean. States 1, 3 and 5 have radial means near 1 nm or less, which suggest that they
  occur close to the pore centers. States 2, 4 and 6 occur well within the tail region of the membrane.
  (c) Using the transition probabilities (p), we estimated the average time spent within each 
  of the states using 1000 simulations of a Bernoulli process where we continued to draw until transitioning
  out of the state with probability 1 - p. Of the prevelent states, states 1 and 
  2 exhbit the shortest dwell times alongside their high covariances. States 3 and 6 have the longest dwell
  times where methanol appears to be trapped near the pore center.   
  }\label{fig:common_states_MET}
  \end{figure}

  Our speculation so far is based purely on the parameters of our model compared to 
  a single methanol trajectory.
  \begin{itemize}
    \item While we may be able to further corroborate our arguments by analyzing more
    individual trajectories, it is more useful to correlate the state behavior with
    solute-membrane interactions.
  \end{itemize}

  The types of hydrogen bonding undergone by solutes is dependent on radial
  location (see Figure~\ref{fig:hbond_pi_chart}).
  \begin{itemize}
    \item States with radial means near 1 nm or less (states 1, 3 and 5) typcially
    hydrogen bond with the monomer carboxylate groups and ether linkages.
    \item States with radial means greater than 1 nm (states 2, 4, 6 and 7) tend 
    to hydrogen bond with the monomer tails.
    \item In the past we've shown how there is a gradual transition from the
  	hydrophilic to hydrophobic region in these membranes, meaning the monomer
  	location is not tightly bound in $r$.
  	\item Therefore, there are instances of hydrogen bond interactions that one
  	might consider uncharacteristic based on the state's radial mean.
  \end{itemize}

  States which exhibit longer dwell times have longer hydrogen bond lifetimes.
  \begin{itemize}
    \item States 3--7 have significantly longer hydrogen bond lifetimes than states
    1 and 2. 
    \item It makes physical and mathematical sense that states 1 and 2 should not 
    stay trapped by hydrogen bonding for long periods because they typically take large
    hops as communicated by their large covariances.
    \item In the tail states, hydrogen bond lifetimes appear to increase with 
    their dwell times.
    \item However, in the pores, this is not necessarily a linear relationship. 
    \item For example, state 5 has the lowest hydrogen bond lifetime of states 3--6
    despite having the second longest dwell times. 
  \end{itemize}

  The local density experienced by solutes may have an impact on the length of solute 
  entrapment.
  \begin{itemize}
    \item On average, solutes in state 5 are surrounded by an above average number of
    particles. 
    \item Based on its radial location and preference towards hydrogen bonding
    with the ether linkages, it is reasonable to assume that while in state 5, methanol
    spends most of its time just outside the pore region, sandwiched between the 
    monomer head groups. 
    \item Although state 5 hydrogen bonds break more frequently than one might expect for 
    a state with long dwell times, it is possible that the solute's dense surroundings
    prevent it from wandering off before the hydrogen bonding interaction can be
    recovered.
    \item This may explain the behavior of methanol in state 5 of the trajectory in
    Figure~\ref{fig:mechanism_map} where the solute's center of mass appears to move
    more freely about its mean than other trapped states.
    \item The data suggests that methanol molecules in state 3 lie closer to the pore
    center than state 5 because they tend to hydrogen bond most frequently with 
    the carboxylate groups.
    \item Methanol molecules which experience state 3 are subject to a lower density 
    local environment and rely more heavily on hydrogen bonding interactions, 
    evidenced by their longer lifetimes, to stay trapped.
  \end{itemize}

  States describing behavior in the tail region show a direction relationship between 
  hydrogen bond lifetimes and local density.
  \begin{itemize}
    \item Methanol molecules which experience state 2 are in a low density region of the
    tails where they can easily break free of hydrogen bonds.
    \item Methanol molecules in states 4 and 6 have hydrogen bond lifetimes that are similar
    to the mean dwell times, which suggests
    \item Finally, methanol molecules trapped in state 7 are in the densest region of all
    states studied. It is likely difficult for methanol to esacpe these high density regions
    and its position is further stabilized via hydrogen bonding.
  \end{itemize}

@@

  %BJC3: this figure is not even close to done
  %BJC3: I've tried a couple different things to look at the hbonds. The other was 
  % how often they are hydrogen bonded in general versus not hydrogen bond. They all
  % hydrogen bond about the same amount. So I think hbond lifetimes may be the key. 
  % I need to implement this first to test that hypothesis.
  %BJC4: I didn't put sodium ion assocaition in this because it's infrequent. 
  %MRS5: is it correlated with specific states?
  %BJC4: If I compare to acetic acid or urea, I could add it back. I was thinking of splitting
  % the pi chart in half with slices corresponding to hbonds on one side and association on the other
  \begin{figure}
  \centering
  \begin{subfigure}{0.54\textwidth}
  \includegraphics[width=\textwidth]{hbond_pi_charts.pdf}
  \caption{}\label{fig:hbond_pi_charts}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{hbond_lifetimes_density_MET.pdf}
  \caption{}\label{fig:hbond_lifetimes_density_MET}
  \end{subfigure}
  \caption{With our clustering approach, methanol exhibits similar state behaviors
  %MRS5: So this is a little counterintuitive, in that the separate states have similar behavior.
  %MRS5: Are there charts that show what is _different_ about the states as well?  The previous figure
  %      Shows that they are dynamically different, but this structural pie charge says the existance
  %      of H-bonding is not the reason . . . 
  in all regions of the membrane. In all of the common states described in 
  Figure~\ref{fig:common_states_MET}, methanol, at some point, hydrogen bonds with 
  each region of the LLC monomer to an appreciable degree. However, methanol frequently
  is not hydrogen bonded at all.
  %BJC4: hypothesis yet to be tested. Somehow, I'll incorporate lifetimes into this plot
  %MRS4: lifetimes would be good!
  We can learn more by looking at the hydrogen bond lifetime. In states 2 and 4, solutes
  stay hydrogen bonded for longer periods of time.
  %MRS5: OK, here is some physical insight - local density is what determines trapping. Emphasize the physical mechanisms (wat does and doesn't come out physically because of these analyses) like this more explicitly.
  %MRS5: can you make other graphs like the H-bond graph? like histogram/pie chart of local density
  %MRS5: as a function of state?  What about r as a function of state?  what other ways are there to show
  %MRS5: the characteristics of the states?
  These longer lifetimes exist because, when hydrogen bonded, the local density is higher, preventing the 
  bond from breaking so easily. In states 1 and 3, hbonds are fleeting interactions because
  the surrounding density is low, allowing solutes to move more freely.
  }\label{fig:hbond_pichart}
  \end{figure}

### 2020-04-07 Ben_Manuscripts/thesis/intro.tex
Reworking outline with radial clusters

  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/membrane_separation_size_regimes.pdf}
  \caption{Caption of the future}\label{fig:size_regimes}
  \end{figure}

@@

  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/lc_phases.pdf}
  \caption{}\label{fig:lc_phases}
  \end{figure}

  %BJC: figure TBD here showing each monomer and the phase they form

@@

  This thesis uses the high level of detail afforded by MD simulations in
  order to understand solute transport within LLC membranes with molecular 
  detail. Before diving into atomistic model, we offer a brief survey of 
  the most popular macroscopic treatments in order to highlight how an
  atomistic model may be better suited to understanding how to design LLC
  %TODO

  %MRS3: one issue is that after this paragraph, it's not entirely
  %clear where you are going.  This describes common models, but not
  %how your research is placed in respect to them.
  %BJC3: Added above connecting paragraph

@@

  Membrane permeation is the result of a chemical potential gradient, $\frac{d\mu_i}{dx}$.
  The flux of component $i$ is proportional to this gradient:
  \begin{equation}
    J_i = -L_i \frac{d\mu_i}{dx}
  \end{equation}
  where $L_i$ is a coefficient of proportionality.~\cite{wijmans_solution-diffusion_1995}
  $L_i$ is an intrinsic materials property dependent on ... % Wijman's doesn't say. I don't want to speculate. Get book he referenced from library
      %MRS2: mention what L depends on molecularly (even if uncalculatable)
  Chemical potential gradients can be induced by concentration, pressure, temperature
  and electromotive forces. We will limit our discussion to concentration and pressure
  gradients because these are primarily forces used for separations with the types of 
  membranes we will be studying.

>>>>

  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/membrane_separation_size_regimes.pdf}
  \caption{Choosing a membrane separation technology is entirely dependent on the
  size of the particle to be removed. The diagram provides examples of the types 
  of particles that each class of membranes is able to remove. Microfiltration (MF) 
  membranes contain pores with diameters ranging from 100-10000 nm and can separate
  large particles like bacteria and protozoa.~\cite{ma_functionalized_2014}
  Ultrafiltration (UF) membranes have pore diameters of about 5-500 nm and are 
  useful for the separation of sugars, proteins, viruses and colloidal 
  materials.~\cite{yanful_ultrafiltration_2009} Nanofiltration (NF) membranes have
  pores on the order of 1 nm in diameter and can be used for the separation of 
  small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
  Reverse osmosis (RO) membranes are dense amorphous polymers with tortuous 
  pathways formed by collapsed polymer networks. Their dense polymer architecture 
  rejects all solutes to a high degree except for water making them useful for 
  separating hydrated ions from water.~\cite{warsinger_review_2018}
  }\label{fig:size_regimes}
  \end{figure}

@@

  \begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/lc_phases.pdf}
  \caption{The amphiphilic character of liquid crystal monomers drives their
  self-assembly into various nanostructured phases with uniform sized pores. 
  The hydrophilic components of the monomers line the aqueous pores, offering
  potential for selective interactions with passing solutes. The monomer on 
  the top left will form the inverted hexagonal (H\textsubscript{II})
  phase which is composed of aligned cylindrical pores packed in a hexagonal 
  array. This geometry is theoretically ideal for high throughput transport. 
  The monomer on the right will for the type I bicontinuous cubic phase
  (Q\textsubscript{I}) in either the Ia3d (gyroid) or Pn3m (diamond) space group.
  Although the Q\textsubscript{I} phase has a much more tortuous network of 
  pores than the H\textsubscript{II}, which should lower its relative permeabily,
  its synthesis is much more practical because it does not require alignment of
  the nanostructured mesophases.
  }\label{fig:lc_phases}
  \end{figure}

@@

  This thesis uses the high level of detail afforded by MD simulations in
  order to understand solute transport within LLC membranes with molecular 
  detail. Before diving into an atomistic model, we offer a brief survey of 
  the most popular macroscopic treatments in order to highlight how a 
  nanoscopic perspective offered by atomistic modeling may be better suited to 
  understanding how to design the next generation of LLC membranes.

@@

  Membrane permeation is the result of a chemical potential gradient, $\frac{d\mu_i}{dx}$.
  The flux of component $i$ is proportional to this gradient:
  \begin{equation}
    J_i = -L_i \frac{d\mu_i}{dx}
  \end{equation}
  where $L_i$ is a coefficient of proportionality.~\cite{wijmans_solution-diffusion_1995}
  $L_i$ is an intrinsic materials property dependent on the chemical makeup of
  the membrane and usually determined via experiment. Chemical potential gradients
  can be induced by concentration, pressure, temperature
  and electromotive forces. We will limit our discussion to concentration and pressure
  gradients because these are primarily forces used for separations with the types of 
  membranes we will be studying.
